# Data preparation

give an outline of the data and major challenges

```{r}
# Load the packages
library(readr)
library(dplyr)
```

```{r}
# Read in the data from the raw folder
#stroke_dt <- read_csv("stroke.csv")
View(stroke)
#stroke_dt <- stroke
#
```

## Cleaning & Understanding Data

```{r}
# Check the dimensions and structure of the data
glimpse(stroke_dt)
str(stroke_dt)
summary(stroke_dt)

```

| using `glimpse(), str(), summary()`following are the observation drawn

1.  Date format has Full week, month and year with century structure

2.  most of the columns which are in num class needs to be changed

3.  for further understanding data cleaning is required

    ```{r}
    stroke_dt$gender <- as.factor(stroke_dt$gender)
    stroke_dt$ever_married <- as.factor(stroke_dt$ever_married)
    stroke_dt$work_type <- as.factor(stroke_dt$work_type)
    stroke_dt$Residence_type <- as.factor(stroke_dt$Residence_type)
    stroke_dt$smoking_status <- as.factor(stroke_dt$smoking_status)
    stroke_dt$bmi <- as.integer(stroke_dt$bmi)
    ```

    To work with values in character class data coercion is required
    Using factor function on these character class values a value can
    be established 

    ```{r}
    # checking the factor result 
    str(stroke_dt)
    summary(stroke_dt)
    ```

    Changing date structure

    ```{r}
    # Check for repeating ids 
    dim(stroke_dt[duplicated(stroke_dt$id),])[1]
    # Since there are no continuous records for same id we will not use date 
    # there is no need for converting date column 
    ```

    Checking for NA

    ```{r}
    colSums(is.na(stroke_dt))
    ```

    Only bmi has NA values. Further consideration is required to make 
    a decision on bmi NA's

    ```{r}
    bmi_na_percentage <- (sum(is.na(stroke_dt$bmi))/length(stroke_dt$bmi)*100)
    percentage_of_Bmi_stroke <- (sum(stroke_dt$stroke                                                                                [is.na(stroke_dt$bmi)])/
                                    sum(stroke_dt$stroke >= 1)*100)
    percentage_of_Bmi_stroke
    bmi_na_percentage
    ```

    Overall NA % in BMI is 3.93%.That holds a 16.06% of instance where stroke
    occurred.
    Repleacing these NA's with mean value is a better option considering their 
    effect on stroke data

    ```{r}
    # Replace the missing values with mean of bmi attribute
    stroke_dt$bmi[is.na(stroke_dt$bmi)] <- mean(stroke_dt$bmi,na.rm = TRUE)
    sum(is.na(stroke_dt$bmi))
    ```

    ```{r}
    summary(stroke_dt)
    
    ```
    writing the completed source file 
```{r}
  write.csv(stroke_dt, file = "stroke.csv")
```
    
    